Issues with Dynamical Measurements Across the Disk

### 2019-02-15 paper/ms.tex
Rewrite abstract

We demonstrate the effect of midplane errors on the computation of stellar
actions. This error is induced by the fact that the local Galactic midplane
varies across the disk. As a result, when we make 6D phase space measurements
of stars that are non-local (i.e. far away), we are observing those stars with
a systematic offset in $z$. We show that this systematic offset induces time
evolution in the observed actions which we interpret as an error. We note that
this error is non-Gaussian and bimodal, with neither of the modes peaking on
the true action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
and thick disk orbits, respectively. Using cosmological zoom-in simulations
from the FIRE collaboration, we show that the midplane can vary by
$\sim100\,\pc$ around the Solar circle, indicating that such an effect is
important for understanding the thin and thick disk. Our discussion is also
directly applicable to variations in the local standard of rest. The variation
of the midplane must be taken into account when performing dynamical modelling
across large regions of the disk accessible with the {\em Gaia} mission.

>>>>

Stellar actions, computed for stars from their 6D phase space measurement and
an assumed Galactic potential, are used to label orbits. In principle, actions
have the attractive quality of being invariant and are thus good labels of a
star's orbit. Of course, inaccurate measurements of the 6D phase space
position and uncertainities in the assumed Galactic potential induce an error
in the computed action. We show that, in addition to these complications, a
systematic bias in the 6D phase space position induces a phase-dependence in
the actions, which we interpret as an error. This error is non-Gaussian and
bimodal, with neither of the modes peaking on the true action value. A
midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is necessary to induce a
$100\,\%$ error in the vertical action $J_z$ for thin and thick disk orbits,
respectively. Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in cosmological zoom-in simulations from
the FIRE collaboration. Thus, current state-of-the-art action calculations ---
which assume the global and local midplanes are the same --- include a
systematic offset in $z$. Variation in the local standard of rest will induce
similar issues. The variation of the midplane must be taken into account when
performing dynamical modelling across large regions of the disk accessible
with the {\em Gaia} mission.
